```{r setup,echo=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
  comment =  "#>", 
  collapse = TRUE,
  fig.align = "center")

options(crayon.enabled = TRUE)
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, which = c("output", "message", "error"))

rgl::setupKnitr(autoprint = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

# Road Extraction Data Requirements {#sec-road-extraction-data-requirements}

The road enhancement method, implemented using the `ALSroads` package requires three inputs:

1.  A lidar point clouds in the format of a collection of `.las` or `.laz` format. Each file must be spatially indexed with a lax files (see also the lidR book [chapter 18.3](https://r-lidar.github.io/lidRbook/spatial-indexing.html#spatial-indexing-files)).
2.  A Digital Terrain Model (DTM) that corresponds to the collection coverage. This DTM can be generated with `lidR`, see also the lidR book [chapter 4](https://r-lidar.github.io/lidRbook/dtm.html).
3.  A pre-existing vectorial road network (lines), typically in a shapefile or geopackage.

A fourth input can strongly enhance the method

4.  A vectorial map (polygons) of the water bodies.

## LiDAR Data {#sec-lidar-data}

Road enhancement using `ALSroads` requires LiDAR data that covers the roads being updated. For use in the `ALSroads` package lidar data must be:

1.  Non-normalized: to preserve the geometry of the landscape
2.  Spatially indexed: to query areas of interest quickly an improve performance
3.  Include intensity values: that are used by the method
4.  In projected coordinate system and in meters: because some parameters are hard coded in meters
5.  Have a density of 5-10 points/m²: less than that is not guaranteed to work, more than that is not useful and will increase the computation time. If the density of the input point cloud is over 10 points/m² it can be decimated on-the-fly using the `filter` argument (see lidR book [chapter 2.1.1](https://r-lidar.github.io/lidRbook/io.html#filter))

```{r, fig.width=5}
library(lidR)
ctg <- readLAScatalog("data/05_data_load/ctg_example/",  filter = "-drop_withheld -keep_random_fraction 0.25")
plot(ctg)
```

## DTM Data {#sec-dtm-data}

Road enhancement using `ALSroads` requires a Digital Terrain Model (DTM) as an input. DTMs, which can broadly be described a an 'image' of the ground, are used for several purposes that induce, stream flow, basin hydrology, and watershed analysis. The `ALSroads` method uses the 1 meter resolution DTM to estimate the slopes and roughness of the terrain. For use in the `ALSroads` package the input DTM must:

1.  Have a resolution of at least 1 m (higher resolution will be downscaled on-the-fly)
2.  Cover the extent of the LiDAR coverage
3.  Match the coordinate system of the LiDAR coverage

```{r, message=FALSE, fig.width=5}
library(raster)
dtm <- raster("data/05_data_load/DTM_new.tif")
plot(ctg)
plot(dtm, col = gray(1:50/50), add = T)
```

## Existing road network {#sec-existing-road-network}

Road enhancement using `ALSroads` updates an existing road network. Therefore, an existing network is a critical input. The method will not enhance the roads that are not on the existing network.

The existing road network must:

1.  Be a simple feature (sf) format (line)
2.  Match the CRS of the LiDAR coverage.

```{r, fig.width=5}
library(sf)
roads <- st_read("data/05_data_load/roads.shp", quiet = TRUE)
plot(ctg)
plot(st_geometry(roads), add = T)
```

## Water bodies (optional) {#sec-water-bodies}

An optional input for road extraction using `ALSroads` is a set of spatial polygons (sf format) of water bodies. An accurate map of water bodies is not necessary strictly speaking to perform the computation but:

1.  It helps to find the road properly by masking lakes that can easily be interpreted as highly drivable areas because they perfectly flat, homogeneous, and located in cleared areas. From a geometrical point-of-view lakes appeare to be the best place to drive a car. Unless the LiDAR point cloud is water classified, there is no way to guess that these flat cleared areas are not passable.
2.  Water bodies are required to correctly identify and update roads that cross rivers using bridges. In a DTM a river is only a gap with strong slopes each edges; which is an impassable obstacle and not a road. Bridges are visible in the point cloud but these points must be classified as bridges, otherwise they have the same status as surrounding point. Geometrically speaking a bridge in the point cloud is just a set of points above the grounds, like trees. Therefore, an un-labeled bridge above a river is technically similar to a tree above the ground.
3.  Providing a map of water bodies enables users to inform the method that the road being updated is known to cross a river. Therefore, the method will allow the passage of a gap (as it appears on the DTM) that would be otherwise considered impassable.
