```{r,echo=FALSE,message=FALSE,warning=FALSE}
r3dDefaults = rgl::r3dDefaults
r3dDefaults$FOV = 50
r3dDefaults$zoom = 0.5

knitr::opts_chunk$set(
  comment =  "#>", 
  collapse = TRUE,
  fig.align = "center")
rgl::setupKnitr(autoprint = TRUE)
```

# Tips and Tricks {#sec-tips-and-tricks}

Following the update of road segments using the `ALSroads`package users may wish to edit data, filter results, or plot outcomes. Here, we provide code snippets intended to provide users with some data interpretation tools. 

## Reconnecing Roads

## Casting Roads to Linestring 

```{r ID_col, rgl = TRUE, warning  = FALSE, fig.height=15, eval = FALSE}
roads <- st_cast(roads, "LINESTRING")
```

## Adding an ID Column

```{r ID_col, rgl = TRUE, warning  = FALSE, fig.height=15, eval = FALSE}
roads <- tibble::rowid_to_column(roads, "NUM")
```

## Selecting Specific Roads

```{r subset_roads, rgl = TRUE, warning  = FALSE, fig.height=15, eval = FALSE}
road_subset <- roads %>%
  dplyr::filter(NUM %in% c(0:10))
```

## Plotting Data 

## Saving Data

